\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{subcaption}

\graphicspath{{figures/}}

\title{Validity-Aware Spectral Diagnostics for Multifractal Random Walk Inference Under Finite Samples}
\author{Yin Jiacheng}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Multiscale fluctuations in complex stochastic systems often exhibit heavy
tails, intermittency, and nonlinear scaling spectra, features that are commonly
interpreted as evidence of multifractality. Under finite observations, however,
this interpretation is intrinsically ambiguous: stable monofractal processes
may produce apparently regular scaling, heavy-tailed or regime-switching
signals may generate spurious curvature, and high-order empirical moments can
be strongly affected by sample length, scale range, and the choice of
\(q\)-grid. This paper proposes CMIN-SR, a validity-aware spectral diagnostic
framework for assessing multifractal random walk (MRW)-like evidence under
finite samples. Rather than treating neural inference as unconstrained spectrum
regression, CMIN-SR separates empirical spectrum estimation from spectral
interpretation. An observed scaling spectrum is projected onto competing
monofractal and MRW spectral families, and calibrated diagnostic scores are
reported for scaling stability, nonlinear curvature, monofractal
compatibility, MRW compatibility, boundary behavior, and instability.

Controlled analytic-spectrum experiments show that the spectral-geometry
calibrator learns the intended distinction between linear, boundary, and
curved spectra. Across three random seeds, linear monofractal spectra receive
low MRW compatibility and low curvature scores, whereas strongly curved MRW
spectra receive high curvature and MRW-compatibility scores. In representative
runs, the normalized MRW-compatibility score increases from about \(0.13\) for
linear spectra to about \(0.94\) for curved MRW spectra. On raw finite samples,
zeta-alignment reduces monofractal false positives in the documented
\(T=1024\) comparison, lowering fGn MRW compatibility from about \(0.73\) to
\(0.36\). However, deterministic and classical empirical-spectrum estimators,
including aggregate structure functions, MFDFA, and lightweight Haar
wavelet-style controls, do not remove the short-window curvature bottleneck:
MFDFA improves some error and high-\(\lambda^2\) detection metrics, but
projected \(\lambda^2\) remains weakly and inconsistently correlated with the
known intermittency parameter in the tested regime. These findings identify
empirical spectrum estimation and finite-sample curvature recovery, rather
than spectral interpretation alone, as the central bottleneck in short-window
MRW inference. CMIN-SR should therefore be understood as a calibrated
diagnostic framework for multifractal spectral evidence, not as proof of an
MRW data-generating mechanism.
\end{abstract}

\input{sections/01_introduction}
\input{sections/02_framework}
\input{sections/03_methods}
\input{sections/04_experiments}
\input{sections/05_results}
\input{sections/06_failure_analysis}
\input{sections/07_discussion}
\input{sections/08_conclusion}

\bibliographystyle{plain}
\bibliography{references}

\appendix
\input{sections/appendix_ablation_notes}

\end{document}
